---
title: "Final Project"
subtitle: "STAT 230 Spring 2025"
author: "Alex Gober"
format:
  html:
    embed-resources: true

Below I've created a histogram, box plot, and scatterplot of my data
```{r}

ggplot(Mental_Health_Care_in_the_Last_4_Weeks, aes(x = Value)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Mental Health Value in the Last 4 Weeks",
       x = "Value (e.g., Days with Poor Mental Health)",
       y = "Frequency") +
  theme_minimal()
  
```

```{r}

  ggplot(Mental_Health_Care_in_the_Last_4_Weeks, aes(x = Group, y = Value, fill = Group)) +
  geom_boxplot() +
  labs(title = "Boxplot of Mental Health Value by Group",
       x = "Group",
       y = "Mental Health Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```

```{r}

ggplot(Mental_Health_Care_in_the_Last_4_Weeks, aes(x = Confidence_Interval, y = Value)) +
  geom_point(color = "blue") +
  labs(title = "Scatterplot of Mental Health Value vs Confidence Interval",
       x = "Confidence_Interval",
       y = "Mental Health Value") +
  theme_minimal()
  
```


Below i've summarized the stats of the dataset
```{r}

summary_stats <- Mental_Health_Care_in_the_Last_4_Weeks %>%
  group_by(Group) %>%
  summarize(
    Mean = mean(Value, na.rm = TRUE),
    SD = sd(Value, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),
    Max = max(Value, na.rm = TRUE)
  )
print(summary_stats)

```
 Group                                          Mean    SD   Min   Max
   <chr>                                         <dbl> <dbl> <dbl> <dbl>
 1 By Age                                         16.0  7.72   1.4  33.9
 2 By Disability status                           23.9 13.2    7.8  49.5
 3 By Education                                   16.4  6.70   5.6  30  
 4 By Gender identity                             26.0 15.2    7.3  62.9
 5 By Presence of Symptoms of Anxiety/Depression  20.0 12.6    2.4  45.7
 6 By Race/Hispanic ethnicity                     15.2  6.72   4.4  32.9
 7 By Sex                                         16.6  7.81   6.7  33.8
 8 By Sexual orientation                          26.6 11.3    8.6  51.9
 9 By State                                       17.4  7.66   3.3  39.3
10 National Estimate                              16.8  6.50   8.7  27.1


Ive made a two correlations below.
```{r}

gender_df <- Mental_Health_Care_in_the_Last_4_Weeks %>%
  filter(Indicator == "Received Counseling or Therapy, Last 4 Weeks", Group == "By Sex") %>%
  mutate(GenderCode = ifelse(Subgroup == "Male", 0, 1))

gender_correlation <- cor.test(gender_df$Value, gender_df$GenderCode, method = "pearson")
print(gender_correlation)

```
	Pearson's product-moment correlation

data:  gender_df$Value and gender_df$GenderCode
t = 26.816, df = 64, p-value < 2.2e-16
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 0.9325116 0.9743221
sample estimates:
      cor 
0.9582655 
```{r}

filtered_data <- Mental_Health_Care_in_the_Last_4_Weeks %>%
  filter(!is.na(Value), !is.na(LowCI), !is.na(HighCI))
  numeric_data <- filtered_data %>%
  select(Value, LowCI, HighCI)
cor_matrix <- cor(numeric_data)
print(round(cor_matrix, 2))


```
       Value LowCI HighCI
Value   1.00  0.99   0.99
LowCI   0.99  1.00   0.94
HighCI  0.99  0.94   1.00


Data indicates that much more women than men will be likely to seek therapy or counseling for mental health. Correlation coefficient was very high (0.96) as it clearly suggests a strong link between being female and receipt of mental health support. This difference isn't attributable to chance-rather, it's a truly evident trend in the numbers. 

Although the graphs help provide overall patterns of the data they don't really show who is taking medications. But if there were to be more women in counseling, chances are pretty high there will be a corresponding number for those on mental health medication. All the other data, such as confidence intervals, just hold up the numbers as truth.

While this data does show off how the data is analyzed, some graphs are much better for clearly showing gender differences in mental health treatment than general histograms or scatterplots. Things such as Bar Charts

